# Packages

## Our packages

Oxinfer developed several packages to work with the OMOP Common Data Model, if you click to any of the package hexstickers you will be redirected to the each package website.

```{r, results='asis', echo=FALSE}
darwin <- c(
  "PatientProfiles", "DrugUtilisation", "CohortSurvival", "omopgenerics",
  "CohortCharacteristics"
)
darwin2 <- c("CodelistGenerator", "visOmopResults", "IncidencePrevalence")
oxford <- c("CohortSymmetry", "OmopSketch", "PhenotypeR")
ohdsi <- c("omock", "CohortConstructor")
element <- function(pkg, org) {
  web <- "https://github.com/{org}/{pkg}/blob/main/man/figures/logo.png?raw=true" |>
    glue::glue()
  if (RCurl::url.exists(web)) {
    web <- '<img src="{web}" alt="{pkg}" style="width:150px;">' |>
      glue::glue()
  } else {
    web <- pkg
  }
  '<a class = "hexsticker" href="https://{org}.github.io/{pkg}/">{web}</a>' |>
    glue::glue()
}
text <- "<div>\n"
repos <- character()
for (el in darwin) {
  text <- paste0(text, element(el, "darwin-eu-dev"), "\n")
  repos <- c(repos, paste0("https://github.com/darwin-eu-dev/", el))
}
for (el in darwin2) text <- paste0(text, element(el, "darwin-eu"), "\n")
for (el in oxford) text <- paste0(text, element(el, "oxford-pharmacoepi"), "\n")
for (el in ohdsi) text <- paste0(text, element(el, "ohdsi"), "\n")
cat(text)
```

## Dashboard

```{r, echo = FALSE}

```

- package name
- repo
- website
- citation
- article
- verison
- Last cran release date
- First cran release date
- checks
- cran
- test coverage
- lifecycle
- reverse dependencies
- maintainer
- open issues
