# OxInfer software

A core activity of OxInfer is building and maintaing research software, particularly focused on supporting analyses of health data mapped to the OMOP Common Data Model. For this we make open-source R packages which are released on CRAN.

### Releases

Below is a summary of CRAN releases of our R packages. As can be seen we have aim for a high frequency of releases, particularly in the early stages of package development. This reflects our aim of practicing continuously delivery.[^summary-1]

[^summary-1]: Always keeping our software product in a releasable state, allowing fast release of features and rapid response to any failures. For more information on what continuous delivery means see <https://martinfowler.com/bliki/ContinuousDelivery.html> and for more on its benefits see <https://dora.dev/capabilities/continuous-delivery/>

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(igraph)
library(ggraph)
library(tidygraph)
library(tools)
library(jsonlite)
library(dplyr)
library(purrr)
library(patchwork)
library(ggplot2)

# db <- tools::CRAN_package_db()
source(here::here("packages", "scrapePackages.R"))
packages <- readr::read_csv(here::here("newsletter", "packages.csv"), show_col_types = FALSE)
pkgs <-  packages |> 
  pull("package_name")
type <-  packages |> 
  pull("organisation")
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
get_cran_release_history <- function(package_name) {
  url <- paste0("https://crandb.r-pkg.org/", package_name, "/all")
 
  releases <- list()
  for(i in seq_along(url)){
  cli::cli_inform("Getting releases for {package_name[i]}")
  releases[[i]] <- fromJSON(url[i]) |>
    pluck("versions") |>
    map_dfr(~ tibble(
      package_name = package_name[i],
      type = type[i],
      version = .$Version,
      date = as.Date(.$Date)
    )) |> 
    arrange(date)
  
  # manual fix
  releases[[i]] <- releases[[i]] |> 
    dplyr::mutate(
      version = dplyr::if_else(
        package_name == "CodelistGenerator" & 
          version == "3.0",
        "3.0.0",
        version))
  
  releases[[i]] <- releases[[i]] |> 
    dplyr::mutate(
      first = as.integer(gsub("^(\\d+)\\..*$", "\\1", version)),
      second = as.integer(gsub("^\\d+\\.(\\d+)\\..*$", "\\1", version)),
      third = as.integer(gsub("^.*\\.([0-9]+)$", "\\1", version))
    ) |> 
    dplyr::mutate(previous_first = lag(first),
                  previous_second = lag(second),
                  previous_third = lag(third)) |> 
    dplyr::mutate(release_type = dplyr::case_when(
      is.na(previous_first) ~ "first release",
      first > previous_first  ~ "major release",
      second > previous_second  ~ "minor release",
      third > previous_third  ~ "patch release",
      .default = NA
    )) |> 
    dplyr::select(package_name,
                  type,
                  version, date, release_type)
  }
  
  dplyr::bind_rows(releases)
  
}


release_hist <- get_cran_release_history(pkgs)
ranges <- release_hist |>
  group_by(package_name, type) |>
  summarise(min_date = min(date), max_date = Sys.Date())

p1 <- release_hist |>
  mutate(package_name = factor(package_name,
                               levels = rev(pkgs))) |> 
  ggplot() +
  geom_segment(data = ranges, 
               aes(x = min_date, xend = max_date,
                   y = package_name,
                   yend = package_name),
               color = "grey90", linewidth = 1.5) +
  geom_point(aes(date, package_name,
                 fill = release_type,
                 size = release_type),
             shape = 21) +
  scale_size_manual(
    values = c("first release" = 3, 
               "patch release" = 1, 
               "minor release" = 2, 
               "major release" = 3)
  ) +
  scale_fill_manual(
    values = c("first release" = "#e41a1c", 
               "patch release" = "#377eb8", 
               "minor release" = "#4daf4a", 
               "major release" = "#984ea3")
  ) +
  theme_bw() +
  ylab("") +
  xlab("") +
  # facet_grid(type~., scales = "free_y") +
  theme(legend.title = element_blank(),
        legend.position = "top",
        panel.spacing = unit(0, "lines"))
p1 <- p1+ theme(
  axis.text.x = element_blank(), 
  axis.title.x = element_blank(),
  axis.ticks.x = element_blank()) 

p2 <- release_hist |> 
  mutate(date_floor = lubridate::floor_date(date, unit = "month")) |> 
  group_by(date_floor) |> 
  tally() |> 
  ggplot() +
  geom_col(aes(date_floor, n), width = 30, colour = "black", fill = "grey70")+
  theme_bw() +
  ylab("") +
  xlab("Releases in month")

p1 / p2 + 
  plot_layout(heights = c(4, 1))
```

### Relationships

The figure below is an attempt to show how our R packages relate to each other. We can see in the centre the core, lower-level, packages. Key imports across packages include `omopgenerics`, which defines core classes and methods for working with data in the OMOP Common Data Model format, and `PatientProfiles`, which provides functionality for identifying and bringing together patient features spread across OMOP Common Data Model tables. In addition, key suggests include `omock`, which creates synthetic datasets, `visOmopResults`, which supports creation of plots and tables to visualise study results. Meanwhile on the periphery are more user-facing, higher-level, packages. For example, these include `OmopSketch`, for summarising a dataset mapped to the OMOP Common Data Model, and `PhenotypeR`, for running diagnostics on them.  

```{r, echo=FALSE, message=FALSE, warning=FALSE}
get_edges <- function(pkg, type) {
  pkg_db <- available.packages(repos = "http://cran.us.r-project.org")
  deps <- package_dependencies(pkg, 
                               db = pkg_db, 
                               which = type, 
                               recursive = FALSE)[[1]]
  internal_deps <- intersect(deps, pkgs)
  
  if (length(internal_deps) > 0) {
    return(data.frame(from = pkg, to = internal_deps, type = type))
  } else {
    return(NULL)
  }
}
plot_dependencies <- function(suggests){
edge_list <- list()
for (pkg in pkgs) {
  edge_list[[paste0(pkg, "_imp")]] <- get_edges(pkg, "Imports")
  if(isTRUE(suggests)){
  edge_list[[paste0(pkg, "_sug")]] <- get_edges(pkg, "Suggests")
  }
}

final_edges <- do.call(rbind, edge_list)
g <- as_tbl_graph(graph_from_data_frame(final_edges, directed = TRUE, vertices = pkgs))
g <- g %>% 
  mutate(
    dependencies = centrality_degree(mode = "in")
  )
ggraph(g, layout = "centrality", centrality = dependencies) + 
  geom_edge_arc(aes(linetype = type, color = type), 
                strength = 0.1, end_cap = circle(4, 'mm'),
                arrow = arrow(length = unit(3, 'mm'), type = "closed")) +
  geom_node_point(aes(fill = dependencies),
                  shape = 21, colour = "black",
                  size = 5) +
  geom_node_text(aes(label = name), 
                 repel = TRUE,         
                 bg.color = "white",   
                 bg.r = 0.15,
                 size = 4) +
  scale_edge_linetype_manual(name = "Dependency type",
                             values = c("Imports" = "solid", 
                                        "Suggests" = "dashed")) +
  scale_edge_color_manual(name = "Dependency type",
                          values = c("Imports" = "black", 
                                     "Suggests" = "grey45")) +
  scale_fill_viridis(name = "Reverse dependencies",
                     option="mako",
                     direction = -1) +
  theme_void() +
  theme(legend.position = "bottom", 
        legend.box = "veritcal")
}
```

::: {.panel-tabset}
### Imports 
```{r, echo=FALSE, message=FALSE, warning=FALSE}
plot_dependencies(suggests = FALSE)
```

### Imports and suggests
```{r, echo=FALSE, message=FALSE, warning=FALSE}
plot_dependencies(suggests = TRUE)
```

:::

### Packages

```{r, results='asis', echo=FALSE, warning=FALSE, message=FALSE}
readr::read_csv(here::here("newsletter", "packages.csv"), show_col_types = FALSE) |>
  dplyr::filter(.data$dev == "ox") |>
  dplyr::select(!"dev") |>
  purrr::pmap(\(package_name, organisation) {
    summarisePackage(pkg = package_name, org = organisation)
  }) |>
  paste0(collapse = "\n\n") |>
  cat()
```
